Read CMEMS netcdf file to R and extract variables

Go to https://data.marine.copernicus.eu/product and select required product. Select variables, study area and time period accordingly.

This example shows monthly means of chlrophyll and disolved oxygen from following product. https://data.marine.copernicus.eu/product/GLOBAL_MULTIYEAR_BGC_001_029/download?dataset=cmems_mod_glo_bgc_my_0.25_P1M-m_202112

```{r}
require(dplyr)
require(readr)
require(here)
library(ggplot2)
library(RNetCDF)


here::i_am("individual_files/sudheera/Read_NC_file_to_R.qmd")
#dir_data <- here::here("Data/")

```

  
 Read netcdf file
 
```{r}
  nct <- open.nc("~/ihw23_proj_tuna/individual_files/sudheera/cmems_mod_glo_bgc_my_0.25_P1M-m_1695268416798.nc", Write=False)
  print.nc(nct)
  
```
  See 
  
  dimensions:
	time = 12 ;
	depth = 1 ;
	latitude = 81 ;
	longitude = 81 ;
  
  variables
  o2 and chl
  
  Read latitude and longitude and create a grid to store variables.
  
```{r}
  lat <- var.get.nc(nct,"latitude")
  lon<- var.get.nc(nct,"longitude")
  latlon <- list(lat=lat,lon=lon)
  latlongrid <- expand.grid(latlon)
```
Next, read variables. see discription to see if there any scaling factor or constant to add.

```{r}
  oxy<-var.get.nc(nct,"o2")
dim(oxy)
```
 Oxygen data has 12 months
 
```{r}
  chl<-var.get.nc(nct,"chl")
  dim(chl)
```
 Get time variable and see the time format. see the discription. 
 NC_DOUBLE time(time) ;
		NC_CHAR time:long_name = "Time (hours since 1950-01-01)" ;
		NC_CHAR time:standard_name = "time" ;
		NC_CHAR time:calendar = "gregorian" ;
		NC_CHAR time:units = "hours since 1950-01-01 00:00:00" ;
		NC_CHAR time:axis = "T" ;
		NC_INT time:_ChunkSizes = 1024 ;
		NC_CHAR time:_CoordinateAxisType = "Time" ;
		NC_DOUBLE time:valid_min = 613980 ;
		NC_DOUBLE time:valid_max = 622020 ;
		
	Time needs to convert.
```{r} 

  t <-var.get.nc(nct,"time")
 print(t)
```
Time needs to convert. as.POSIXlt() can use to get the correct time.
```{r}
  tim <- as.POSIXlt(t*3600,origin="1950-01-01 00:00:00")
  tims <- substring(tim, 1,10)
  tail(tim)
  length(t)
```
Depth variable. Finally we needs to close the nc file. 
```{r}
  
  depth <- var.get.nc(nct,"depth")
  close.nc(nct)
```
  plot the results
```{r}
  library(stars)
  mydata <- stars::read_stars("~/ihw23_proj_tuna/individual_files/sudheera/cmems_mod_glo_bgc_my_0.25_P1M-m_1695268416798.nc")
#> vo, uo,
 str(mydata)
 # plot o2 data
  plot(mydata["o2"], col = hcl.colors(100), axes = TRUE)
  # plot chl data
  plot(mydata["chl"], col = hcl.colors(100), axes = TRUE)
  

```
  
  
  
  
  
  for (i in 1:length(tims)) {
  # sst has 14 depths and 365 days
  # select day i
    #i=1
  latlongrid <- expand.grid(latlon)
  o2 <-oxy[,,i]
  latlongrid$oxy <- c(t(unlist(o2)))
  chla <-chl[,,i]
  latlongrid$chl <- c(t(unlist(chla)))
  
   # save file
  #s.date <- ded <-as.Date(i-1, origin = "2016-01-01")
  save.file <- paste("G:/MLSDM/cmems/chl/",tims[i], ".csv", sep="" )
  write.csv(latlongrid, file=save.file,row.names = F)
  print(tims[i])
  
  } 
```
  
  
  
  